---
title: "Gov 50 Final Project"
author: "Tsion Debebe"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Ideas

-I want to understand how different kinds of factors impact the likeliness of a country experiencing a civil war at least once in a lifetime. 

-The concept of authoritarian regimes is extremely interesting because different democratization levels are usually dependent on certain factors like strength of military, freedom of speech, and natural resource revenue. If I can focus on some of these factors for my project, it would be very interesting.

## Hypothesis

For my project, I want to find the correlation between polity (levels of democracy), population, and oil. I want to study how countries in the Middle East and North African are likely to have high or low democracy levels by analyzing the relationship between polity, popualtion, and natural resource. After analyzing the dataset titled, "civil war", I can hypothesize that countries with higher populations and more oil are less likely to be democratic. My dataset consists of all countries and different variables however, I am soley focused on Middle East/North African countries and oil, population, and polity as factors. Therefore, my outcome variable is polity. I will test my hypothesis by creating a liner regresission and anaylzing visuals that represent the relationship between polity and the different variables.

```{r echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(gapminder)
civil_war <- read_csv("civilwar.csv")

```
## Results 

```{r echo=FALSE}
civil_war<- civil_war |>
  filter(nafrme == 1)

civil_war <- civil_war |>
  mutate(Natural_Resource = if_else(Oil == 1,
                                      "Oil",
                                      "No Oil"))

plot_q0 <- civil_war |>
  group_by(country)|>
  ggplot(aes(y=polity2, x=Natural_Resource)) +
  geom_boxplot()

plot_q1 <- civil_war |>
  group_by(country)|>
  ggplot(aes(y=polity2, x=pop)) +
  geom_point()+
  geom_smooth()


plot_q0
plot_q1

```
The box plot above displays the relationship between Natural Resource (oil) and democracy levels. For instance, Middle Eastern/North African countries with no oil, have a much higher median than countries with oil. The median level of democracy is about -2.7 which means they have more democratization and a higher polity score. 

Additionally, countries with oil have lower levels of democracy. For instance, the box plot above shows us that they have a much lower median at about -8.5 which is extremely low. 

The correlation between natural resource (oil) and democracy could be due to how authoritarian regimes utilize natural resource revenue. For instance, many oil-rich countries in the Middle East are considered authoritarian regimes because they use natural resources to repress citizens and distirbute rents (money) to keep the citizens satisfied. That way, even if citizens are living under an authoritarian regime, they still reap many benefits so they have less of an incentive to revolt against the regime. 


The scatter plot above demonstrates the relationship between population and levels of democracy as well. After analyzing the data, there isn't a very clear correlation but, there seems to be a positive correlation between population and polity levels. 

```{r echo=FALSE}
library(broom)

linearregression <- lm(polity2 ~ pop + Natural_Resource, data = civil_war)|>
broom::tidy()|>
select(term, est=estimate, se=std.error, pval=p.value)
linearregression

knitr::kable(linearregression, digits=6, col.names = c("Variable", "Coefficient Estimate", "Standard Error", "P-Value"))

```


The linear regression above represents the coefficients of oil, natural resource, and population. For the population, the number is extremely small and close to 0 so if we were to round, the number would be zero. However, when looking at how much difference one single person would make, it isn't very significant compared to 1,000 people. When interpreting that coefficient, we can understand thar for every 1 person, there is a 0.0000052 unit increase in democracy levels. 

Additionally, for Natural Resource (oil), the coefficient is -6.428115 which means that for every 1 unit of oil, there is a decrease of -6.428116 units on democracy levels. Moreover, if a country were to go from being a country with oil to no oi, then it would impact the polity score with an average of -6.428

The P-value is less than 0.05 which means that the results from the linear regression table are statistically significant. Under the alpha of 0.05, we can reject the null hypothesis that there is no correlation between polity and natural resou 

The intercept in the table is -1.674970 which represents the case if there was a population of 0 people, the polity score would be -1.675.